A user-extensible and adaptable parser architecture

### 2008-07-30 doc/logparser-results.tex
Updated graph data. No longer discarding data from the first run. Minor corrections. Lots of XXX items added.

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times, the results from the first first run discarded, and the
execution time for the remaining 9 runs averaged.  The results from the
first run were discarded because because the rule ordering will be
sub-optimal compared to subsequent runs --- XXX MAYBE I SHOULD NOT DISCARD
THE FIRST RESULT\@?  Saving results to the database was disabled for the
test runs, as that dominates the run time of the program, and the tests are
aimed at measuring the speed of the parser rather than the speed of the
database and the disks the database is stored on.  The computer used for
test runs was a Dell Optiplex 745, equipped with:

@@

The different distribution of log lines during a mail forwarding loop has
interesting effects on execution time, described in detail in
\sectionref{rule ordering for efficiency}, \sectionref{scalability as the
number of rules rises}, and \sectionref{Caching compiled regexes}.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that the typical parser will have a large number of rules.  There
are \numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files used
when generating the results in this document is \numberOFrulesMINIMUM{},
\numberOFrulesMINIMUMpercentage{} of the full ruleset.  The full ruleset is
larger because the Postfix parser is tested with \numberOFlogFILESall{} log
files, to improve its parsing and ensure it correctly parses logs generated
by later Postfix versions.  A second set of statistics was generated using
the minimum ruleset and compared to the statistics generated using the full
ruleset: the percentage execution increase when using the full ruleset
instead of the minimal ruleset for normal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.

@@

This section began with parser scalability, showing the linear relationship
between execution time and input size.  Demonstrates the effect of rule
ordering on execution time, and the unexpected consequences of specific
inputs.  The necessity of caching compiled \regexes{} is attested to by the
third group of graphs, where the difference between caching and not caching
is staggering.  The penultimate section contains a breakdown of the rule
hits accumulated during a single test run of the parser.  Miscellaneous
graphs expected to be useful are collected in the final section.

>>>>

When generating the timing data used in this section, \numberOFlogFILES{}
log files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times and the parsing times averaged.  Saving results to the
database was disabled for the test runs, as that dominates the run time of
the program, and the tests are aimed at measuring the speed of the parser
rather than the speed of the database and the disks the database is stored
on.  The computer used for test runs was a Dell Optiplex 745, equipped
with:

@@

The different distribution of log lines during a mail forwarding loop has
interesting effects on parsing time, described in detail in
\sectionref{rule ordering for efficiency}, \sectionref{scalability as the
number of rules rises}, and \sectionref{Caching compiled regexes}.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that the typical parser will have a large number of rules.  There
are \numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files used
when generating the results in this document is \numberOFrulesMINIMUM{},
\numberOFrulesMINIMUMpercentage{} of the full ruleset.  The full ruleset is
larger because the Postfix parser is tested with \numberOFlogFILESall{} log
files, to improve its parsing and ensure it correctly parses logs generated
by later Postfix versions.  A second set of statistics was generated using
the minimum ruleset and compared to the statistics generated using the full
ruleset: the percentage parsing time increase when using the full ruleset
instead of the minimal ruleset for normal, shuffled and reversed orderings
is shown in \graphref{Percentage parsing time increase of maximum ruleset
over minimum ruleset}.

@@

This section began with parser scalability, showing the linear relationship
between parsing time and input size.  Demonstrates the effect of rule
ordering on parsing time, and the unexpected consequences of specific
inputs.  The necessity of caching compiled \regexes{} is attested to by the
third group of graphs, where the difference between caching and not caching
is staggering.  The penultimate section contains a breakdown of the rule
hits accumulated during a single test run of the parser.  Miscellaneous
graphs expected to be useful are collected in the final section.

### 2008-07-30 papers/sgai-2008/paper.tex
Updated graph data. No longer discarding data from the first run. Minor corrections. Lots of XXX items added.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It provides the link between actions and rules,
allowing either to be improved independently of the other.  The framework
is the core of the architecture and is deliberately quite simple: the rules
deal with the varying log lines, and the actions deal with the intricacies
and complications encountered when parsing.

@@

@@

    \item [program] The Postfix program whose log lines the regex should be
        applied to.  This avoids needlessly trying regexes that will not
        match the log line, or worse, might match unintentionally.  Rules
        whose program is \texttt{*} will be tried against any log lines not
        parsed by program specific rules, allowing generic rules to be
        written.

@@

Adding new rules is simple: simply define all the required attributes and
append the new rule to wherever the existing rules are stored.  The
software developed for this project includes a utility based on
SLCT~\cite{slct-paper} to aid in producing regexes to parse previously
unparsed inputs.  Storing the rules separately allows other approaches to
rule management to be used, e.g.\ instead of manually adding new rules when
unparsed log lines are seen, machine learning techniques could be used to
automatically generate new rules, based on the existing rules and input
lines.  If this approach was taken the choice of machine learning technique
would be constrained by the size of typical data sets: the
\numberOFlogFILES{} log files used to generate statistics contain
\numberOFlogLINEShuman{} log lines.  Techniques which require the full data
set when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques which automatically
determine which inputs from the training set are valuable and which inputs
can be discarded would probably reduce the data required to a manageable
size.  This avenue of research and development has not been pursued by the
author, but could easily be independently undertaken.

@@

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input line, and regexes which consume
large amounts of CPU time during matching.  Defining a regex which matches
every input line is trivial: \verb!/^/! will match the start of the input
line; every input line is guaranteed to start, so this regex would match
every input line.  In general a regex will not consume a large amount of
CPU time when the match is successful; usually CPU time is consumed by a
regex with a lot of alteration and variable quantifiers failing to match.
This topic is beyond the scope of this paper, see
\cite{mastering-regular-expressions} for in-depth discussion.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse very large log files.  The server which generated the log files used
in testing this parser accepts approximately 10,000 mails for 700 users a
day.  Median log file size is 50MB, containing 285,000 log lines --- large
scale mail servers would have much larger log files.  When generating the
timing data used in this section, \numberOFlogFILES{} log files (totaling
10.08 GB, \numberOFlogLINEShuman{} log lines) were each parsed 10 times,
the results from the first first run discarded, and the execution time for
the remaining 9 runs averaged.  Saving results to the database was disabled
for the test runs, as that dominates the run time of the program, and the
tests are aimed at measuring the speed of the parser rather than the speed
of the database and the disks the database is stored on.  The computer used
for test runs was a Dell Optiplex 745:

\begin{tabular}[]{ll}

    CPU         & One dual core 2.40GHz Intel\textregistered{}
                    Core\texttrademark{}2 CPU, \\
                & with 32KB L1 cache and 4MB L2 cache. \\
    RAM         & 2GB 667 MHz DDR RAM\@. \\
    Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk. \\

\end{tabular}

Parsing all \numberOFlogFILES{} log files in one run took
\input{build/timing-run-include.tex}.

The computer was dedicated to the task of gathering statistics from test
runs, and was not used for any other purpose while test runs were ongoing;
any services not necessary for running the tests were disabled.

@@

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage execution increase when using the full ruleset instead of the
minimal ruleset for normal, shuffled and reversed orderings is shown in
\graphref{Percentage execution time increase of maximum ruleset over
minimum ruleset}.  \showgraph{build/plot-normal-vs-smaller}{Percentage
execution time increase of maximum ruleset over minimum ruleset}{Percentage
execution time increase of maximum ruleset over minimum ruleset}

It is clear from \graphref{Percentage execution time increase of maximum
ruleset over minimum ruleset} that the increased number of rules has a
noticeable performance impact with reverse ordering, and to a lesser extent
with shuffled ordering.  The normal ordering (where are the most frequently
matching rules are tried first) shows a mean increase of
\input{build/normal-vs-smaller-percentage-increase.tex} in execution time
for a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.
These results show that the architecture scales very well as the number of
rules increases, and that sorting the rules is an important optimisation
enabling this.

@@

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly.  The parser
issues warnings when it detects any errors, some of which may alert the
user to a problem.  There should be few or no warnings when parsing, and
when finished parsing the state table should only contain entries for mails
which had yet to be delivered when the log files ended, or were accepted
before the log files began.  A sample of 6000 log lines was parsed with all
debugging options enabled, resulting in 167,448 lines of output.  All
167,448 lines were examined in conjunction with the input log file and a
dump of the resulting database, verifying that for each of the input lines
the parser used the correct rule and executed the correct action, which in
turn produced the correct result and inserted the correct data in the
database.  The log segment produced 4 warnings (about mails which had
started before the log file segment), 10 mails remaining in the state
tables, and 1625 connections correctly entered in the database.  Given the
evidence detailed above, the author is confident that the false positive
rate when reconstructing a mail is zero.

>>>>

XXX DISCUSS COMPLEXITY\@?  EVEN JUST MENTION THAT COMPLEXITY RISES WHEN
INCREMENTAL PARSING IS USED, BUT THAT EXTERNAL TOOLS COULD BE USED TO DEAL
WITH IT, AND THAT IT IS ENTIRELY DEPENDANT ON THE RULESET USED\@?

XXX MAYBE MERGE THE AUTHOR SECTIONS IF DESPERATE FOR SPACE\@.

XXX REVIEWER SUGGESTED COMPARING TO EXISTING WORK, BUT I DID NOT FIND ANY
SIMILAR EXISTING WORK\@.  SHOULD SEARCH AGAIN\@.

XXX OVERLAP DETECTION\@: L1 n L2 = E* --- ((E* --- L1) u (E* --- L2)).
PERL REGEXES ARE AT LEAST AS POWERFUL AS A PUSHDOWN AUTOMATA --- SEE
CHOMSKY'S HIERARCHY TO MAKE SURE\@.  CONTAINMENT CANNOT BE DETERMINED FOR
CFGs, INTERSECTION IS POSSIBLE\@.

@@

The framework takes care of miscellaneous support functions and low level
details of parsing, freeing the authors of actions to concentrate on
writing productive code.  It provides the link between actions and rules,
allowing either to be improved independently of the other.  The framework
is the core of the architecture and is deliberately quite simple: the rules
deal with the varying log lines, and the actions deal with the intricacies
and complications encountered when parsing.  XXX MAYBE REMOVE FIRST TWO
SENTENCES\@.

@@

XXX CARL WONDERS IF THE NUMBERS HERE ARE TYPICAL\@?  ALSO SAY SOMETHING
LIKE REJECTION AND IGNORE DO NOT IMPLY UNGRAMMATICAL\@.

@@

    \item [program] The Postfix program whose log lines the regex should be
        applied to.  This avoids needlessly trying regexes that will not
        match the log line, or worse, might match unintentionally.  Rules
        whose program is ``\texttt{*}'' will be tried against any log lines
        not parsed by program specific rules, allowing generic rules to be
        written.  This is a literal ``\texttt{*}'', not a wildcard or
        regex.

@@

Adding new rules is simple: simply define all the required attributes and
append the new rule to wherever the existing rules are stored.  The
software developed for this project includes a utility based on
SLCT~\cite{slct-paper} to aid in producing regexes to parse previously
unparsed inputs.  Storing the rules separately allows other approaches to
rule management to be used, e.g.\ instead of manually adding new rules when
unparsed log lines are seen, machine learning techniques could be used to
automatically generate new rules, based on the existing rules and input
lines.  If this approach was taken the choice of machine learning technique
would be constrained by the size of typical data sets: the
\numberOFlogFILES{} log files used to generate statistics contain
\numberOFlogLINEShuman{} log lines.  Techniques which require the full data
set when training would be impractical; Instance Based
Learning~\cite{instance-based-learning} techniques which automatically
determine which inputs from the training set are valuable and which inputs
can be discarded would probably reduce the data required to a manageable
size.  This avenue of research and development has not been pursued by the
authors, but could easily be independently undertaken.

@@

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input line, and regexes which consume
large amounts of CPU time during matching.  Defining a regex which matches
every input line is trivial: \verb!/^/! will match the start of every input
line.  In general a regex will not consume a large amount of CPU time when
the match is successful; usually CPU time is consumed by a regex with a lot
of alteration and variable quantifiers failing to match.  This topic is
beyond the scope of this paper, see~\cite{mastering-regular-expressions}
for in-depth discussion.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
parse very large log files.  The server which generated the log files used
in testing this parser accepts approximately 10,000 mails for 700 users a
day.  Median log file size is 50MB, containing 285,000 log lines --- large
scale mail servers would have much larger log files.  When generating the
timing data used in this section, \numberOFlogFILES{} log files (totaling
10.08 GB, \numberOFlogLINEShuman{} log lines) were each parsed 10 times and
the parsing times averaged.  Saving results to the database was disabled
for the test runs, as that dominates the run time of the program, and the
tests are aimed at measuring the speed of the parser rather than the speed
of the database and the disks the database is stored on.  The computer used
for test runs was a Dell Optiplex 745:

\begin{tabular}[]{ll}
    \hline
    \noalign{\smallskip}
    XXX         & XXX NEED LABELS\@.                                \\
    \noalign{\smallskip}
    \hline
    \noalign{\smallskip}
    CPU         & One dual core 2.40GHz Intel\textregistered{}
                    Core\texttrademark{}2 CPU,                      \\
                & with 32KB L1 cache and 4MB L2 cache.              \\
    RAM         & 2GB 667 MHz DDR RAM\@.                            \\
    Hard disk   & One Seagate Barracuda 7200 RPM 250GB SATA disk.   \\
    \noalign{\smallskip}
    \hline
    \noalign{\smallskip}
\end{tabular}

Parsing all \numberOFlogFILES{} log files in one run took
\input{build/timing-run-include.tex}.  The computer was dedicated to the
task of gathering statistics from test runs, and was not used for any other
purpose while test runs were ongoing; any services not necessary for
running the tests were disabled.

@@

XXX MAYBE REDUCE THIS\@?

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage parsing time increase when using the full ruleset instead of the
minimal ruleset for normal, shuffled and reversed orderings is shown in
\graphref{Percentage parsing time increase of maximum ruleset over minimum
ruleset}.  \showgraph{build/plot-normal-vs-smaller}{Percentage parsing time
increase of maximum ruleset over minimum ruleset}{Percentage parsing time
increase of maximum ruleset over minimum ruleset}

It is clear from \graphref{Percentage parsing time increase of maximum
ruleset over minimum ruleset} that the increased number of rules has a
noticeable performance impact with reverse ordering, and to a lesser extent
with shuffled ordering.  The normal ordering (where are the most frequently
matching rules are tried first) shows a mean increase of
\input{build/normal-vs-smaller-percentage-increase.tex} in parsing time for
a \numberOFrulesMAXIMUMpercentage{} increase in the number of rules.  These
results show that the architecture scales very well as the number of rules
increases, and that sorting the rules is an important optimisation enabling
this.

@@

Coverage of mails is much more difficult to determine accurately than
coverage of log lines.  The parser can dump its state tables in a human
readable form; examining these tables with reference to the log files is
the best way to detect mails which were not handled properly.  The parser
issues warnings when it detects any errors, some of which may alert the
user to a problem.  There should be few or no warnings when parsing, and
when finished parsing the state table should only contain entries for mails
which had yet to be delivered when the log files ended, or were accepted
before the log files began.  A sample of 6000 log lines was parsed with all
debugging options enabled, resulting in 167,448 lines of output.  All
167,448 lines were examined in conjunction with the input log file and a
dump of the resulting database, verifying that for each of the input lines
the parser used the correct rule and executed the correct action, which in
turn produced the correct result and inserted the correct data in the
database.  The log segment produced 4 warnings (about mails which had
started before the log file segment), 10 mails remaining in the state
tables, and 1625 connections correctly entered in the database.  Given the
evidence detailed above, the authors are confident that the false positive
rate when reconstructing a mail is zero.
